---
title: "Lab 10: Filtering Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
theme_set(theme_minimal())
```

## Pick a dataset

Today you will again have a choice of three datasets to work with. You can
choose either to investigate the daily team ELO ratings in the NBA:

```{r, message=FALSE}
nba <- read_csv("https://statsmaths.github.io/stat_data/nba_elo_daily.csv")
```

## Instructions

This lab is similar to the last. You'll pick a dataset and then generate an
analysis of that data using data visualizations. The difference here is that
instead of a single plot, I would like you to construct 3-4 plots that
together tell a linear story. Each plot should be separated by a sentence or
two describing what the viewer should take away from the plot. Try to keep the
plots similar in some way; perhaps each simply highlights a different subset
of the data but has the same underlying layers.

Notice that each of the datasets for today are larger than you will probably
be able to use. You will need to filter the data to a particular team, year,
subset of curse words. You may further filter the data to highlight an even
smaller subset of the data within each plot.

## Analysis

```{r}
nba_new <- filter(nba, team %in% c("BOS","NYK", "CHS", "TRH", "PRO", "DTF", "PIT", "WSC", "CLR", "STB"))
```

The graph below narrows the data field to focus in on 10 NBA teams. It highlights the relationship between each team and their ELO over the past seasons.

```{r}
ggplot(nba_new, aes(team, elo)) +
  geom_point() +
  geom_point(aes(color = season))
```

The following graph highlights the first ELO of each team (in the 1947 season). 

```{r}
ggplot(nba_new, aes(team, elo)) +
  geom_point() +
  geom_point(aes(color = season)) +
  annotate("point", x = "BOS", y = 1294.846, color = "red") +
  annotate("point", x = "CHS", y = 1309.652, color = "red") +
  annotate("point", x = "CLR", y = 1307.123, color = "red") +
  annotate("point", x = "DTF", y = 1279.619, color = "red") +
  annotate("point", x = "NYK", y = 1306.723, color = "red") +
  annotate("point", x = "PIT", y = 1295.309, color = "red") +
  annotate("point", x = "PRO", y = 1305.154, color = "red") +
  annotate("point", x = "STB", y = 1304.691, color = "red") +
  annotate("point", x = "TRH", y = 1293.277, color = "red") +
  annotate("point", x = "WSC", y = 1320.381, color = "red")
```

The final graph below shows the most recent ELO of the Knicks and Celtics (in the 2017-18 season). The other eight teams are not highlighted because there is no recent (2017/2018) information provided within the data set.

```{r}
ggplot(nba_new, aes(team, elo)) +
  geom_point() +
  geom_point(aes(color = season)) +
  annotate("point", x = "BOS", y = 1551.458, color = "red") +
  annotate("point", x = "NYK", y = 1418.374, color = "red")
```


