---
title: "Lab 20: Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8.5)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(viridis)
library(forcats)
library(smodels)
theme_set(theme_minimal())
```

## Instructions

Below you will find several empty R code scripts and answer prompts. Your task
is to fill in the required code snippets and answer the corresponding
questions.

## Diamonds Data - Simple Regression

Today, we will start by looking at a dataset of diamond prices:

```{r}
diamonds <- read_csv("https://statsmaths.github.io/stat_data/diamonds.csv")
```

Fit a linear regression predicting the diamond price as a function of
its weight (carats). Print out the regression table without any confidence
intervals.

```{r}
model <- lm_basic(price ~ 1 + carat, data = diamonds)
reg_table(model)
```

How would you interpret the slope coefficient in this model?

**Answer**: The price of the diamond will increase by $7756 for each carat size increased.


Next, show the same regression as a scatter plot with a smoothing line:

```{r}
qplot(price, carat, data = diamonds) + 
  geom_smooth(method = "lm")
```

Now, print the same model as in the previous question but add a 95%
confidence interval.

```{r}
model1 <- lm_basic(price ~ 1 + carat, data = diamonds)
reg_table(model1, level = 0.95)
```

What range of values does the confidence interval give for the intercept
in this model?

**Answer**: Range: -2282 to -2231

What range of values does the confidence interval give for the slope
in this model?

**Answer**: Range: 7729 to 7784

Would you be surprised if we collected another set of diamonds, in exactly the
same way as before, and found a slope estimate equal to 5500?

**Answer**: Yes

Would you be surprised if we collected another set of diamonds, in exactly the
same way as before, and found a slope estimate equal to 7500?

**Answer**: Yes

The biggest use of confidence intervals, at least for us, is to
determine whether the confidence interval for a slope contains only values
that are all negative, all positive, or a mix of both.

Why might this be important? Let's look at our diamond example. Describe the
meaning of the slope in this linear regression:

**Answer**: If the confidence interval contains 0 (aka has positive and negative numbers in it) then it does not count as strong evidence. If it only contains negative (or only contains positive) numbers then it can be refered to as 'strong evidence'.

Therefore, if the confidence interval has only positive values, we can
fairly confidently say is that there is a positive relationship between
diamond weight and its price.

Now, fit a regression model on the diamonds dataset, predicting the price as
a function of depth. Compute the regression table using level equal to .95:

```{r}
model2 <- lm_basic(price ~ 1 + depth, data = diamonds)
reg_table(model2, level = 0.95)
```

What does the model predict will be the price of a diamond with a depth of
60?

**Answer**: Intercept - (depth)*(change in depth)
5763 - (60*29.65) = 4023

Describe in words the estimate of the slope in terms of the dataset.

**Answer**: For a given depth, you would subtract 29.65 from the average price.

Using the regression table from the previous question, does the
confidence interval contain only values that are positive, only values that
are negative, or values that are both positive and negative?

**Answer**: Intercept interval only contains postive numbers.

Interpret this in words (your words, not just the formal definition).

**Answer**: Since the intercept interval only contains positive numbers, there is a positive correlation between the price and depth (as depth increases, price increases).

## Diamonds Data - Multiple Linear Regression

While we lose the nice graphical summary, it is possible to build linear
models with more than one variable and an intercept. In the diamond example,
for instance, we can fit a model that has both depth and carat in it:

```{r}
model3 <- lm_basic(price ~ 1 + carat + depth, data = diamonds)
reg_table(model3, level = 0.95)
```

Notice that switching the order of the inputs does not change the estimates
or confidence intervals; only the order in the output is modified, but this
is inconsequential:

```{r}
model4 <- lm_basic(price ~ 1 + depth + carat, data = diamonds)
reg_table(model4, level = 0.95)
```

Using the model in the previous question, what do you expect will be the
price of a diamond with a depth of 50 that weights 2 carats?

**Answer**: 4045.3 - (50 * 102.2) + (2 * 7765.1) = $14,465.50

Are both of the slopes in the model statistically significant? What are
their signs?

**Answer**: Both are statistically significant (depth is negative, carat is positive)

Use the `add_prediction` function to add predictions from this model
back into the diamonds dataset.

```{r}
model5 <- lm_basic(price ~ 1 + depth + carat, data = diamonds)
diamonds <- add_prediction(diamonds, model5)
```

What is the largest positive residual (you can either use the `max` function
or sort the data in the data viewer)?

**Answer**: Largest positive residual = 12683.655

Interpreting the meaning of the slopes in a regression model with
multiple variables changes slightly from the version with a single slope.
Each slope measures the change in the response give a change in its
corresponding value, when holding the other variables in the model fixed.
This last part is actually important and we will see some examples that at
first seem counterintuitive that make sense once we remember this caveat.

Describe the meaning the slopes, in your own words, in the regression model
above.

**Answer**: The slopes describe the difference of the intercept as each variable increases or decreases.

Notice that we can do some comparisons quickly with the slope estimates.
For example, if we have two diamonds with the same depth where one is 1
carat in weight and the other is 2 carats in weight, how much more do you
expect the heavier diamond to cost?

**Answer**: $7765.10 more

